Finite Time Analysis of Linear Stochastic Approximation

### 2017-05-06 instance.tex
initial version

\section{Instance Dependent Bounds}
In this section, we derive instance dependent bounds for $L_t$ in \eqref{eq:exploss} which is valid for a fixed $P$ and in the next section, explore the question of deriving uniform bounds for all $P\in \P$ (where $\P$ is a class of distributions over $\C^{\ncn}$). We first present the results and then follow it up with a discussion.
\begin{lemma}\label{lm:pd}
Let $P$ be a positive definite distribution over $\C^{\ncn}$, then there exists an $\alpha_P$ such that $\rhod{P}>0$ and $\rhos{P}>0,~\forall \alpha \in (0,\alpha_P)$.
\end{lemma}
\begin{theorem}\label{th:pdrate}
Let $P$ in \Cref{lsadist} be a positive definite distribution over $\C^{\ncn}$. Then
\begin{align}
\EE{\norm{\thh_t-\theta^*}^2}
\leq
\left(1+\frac2{\alpha\rhod{P}}\right)\frac{1}{\alpha \rhos{P}}
\left(\frac{\norm{\theta_0-\theta^*}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\theta_0-\theta^*}}{t+1} \right)\,.
%\left(\frac{\norm{\theta_0-\theta^*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\theta_0-\theta^*})}{t+1} \right)\,.
\end{align}

@@

\begin{lemma}\label{lm:simtran}[Similarity Transformation]
Let $P^{AS}\in\P$ be asymptotically stable, then there exists a $U\in \gln$ such that $P^{AS}_U$ is positive definite.
\end{lemma}

@@

Let $P$ be as in \Cref{lsadist}. Define $\gamma_t\eqdef U^{-1}\theta_t,\,,\gamma^*\eqdef U^{-1}\theta^*$, and suppose if $P_U$ is positive definite, then

@@

\subsection{Results Discussion}
\textbf{Error Recursion and Contraction Maps:}

@@

Notice that \eqref{eq:err} places our definitions of $\rhod{P}$ and $\rhos{P}$ in \Cref{def:contract}. Here the subscripts $s$ and $d$ stand for \emph{stochastic} and \emph{deterministic}, since they are associated with the spectral radius of stochastic and deterministic matrices respectively.\\
\textbf{Positive Definiteness:} The first result in \Cref{lm:pd} shows if $P$ is PD, then there exists a stepsize for which $\rhos{P}, \rhod{P}>0$ which in turn implies that the spectral radii in \eqref{eq:spectralrand} are less than $1$.
\textbf{Similarity Transformation:} 

>>>>

\section{Instance Dependent Bounds}
In this section, we derive instance dependent bounds for $L_t$ in \eqref{eq:exploss} which is valid for a fixed $P$ and in the next section, explore the question of deriving uniform bounds for all $P\in \P$ (where $\P$ is a class of distributions over $\C^{\ncn}$). We first present the results and then follow it up with a discussion.
\begin{lemma}\label{lm:pd}
Let $P$ be a positive definite distribution over $\C^{\ncn}$, then there exists an $\alpha_P$ such that $\rhod{P}>0$ and $\rhos{P}>0,~\forall \alpha \in (0,\alpha_P)$.
\end{lemma}
\begin{proof}
\begin{align*}
\rhos{P}&=\inf_{x:\norm{x}=1}x^* (A_P+A_P^*)x -\alpha x^*\EE{A_t^* A_t} x\\
&=\inf_{x:\norm{x}=1}x^* (A_P+A_P^*)x -\alpha x^* A^*_P A_P -\alpha x^* \EE{M_t^* M_t} x
&\geq \lambda_{\min}(A^*+A)-\alpha \norm{A_P}^2-\sigma^2_P
\end{align*}
The proof is complete by choosing $\alpha_P<\frac{\lambda_{\min}(A^*+A)}{\norm{A_P}^2+\sigma^2_P}$
\end{proof}

\begin{lemma}\label{lm:simtran}[Similarity Transformation]
Let $P^{AS}$ be asymptotically stable, then there exists a $U\in \gln$ such that $P^{AS}_U$ is positive definite.
\end{lemma}

\begin{theorem}\label{th:pdrate}
Let $P$ in \Cref{assmp:lsa}-~\ref{dist} be a positive definite distribution over $\C^{\ncn}$. Then
\begin{align}
\EE{\norm{\thh_t-\theta^*}^2}
\leq
\left(1+\frac2{\alpha\rhod{P}}\right)\frac{1}{\alpha \rhos{P}}
\left(\frac{\norm{\theta_0-\theta^*}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\theta_0-\theta^*}}{t+1} \right)\,.
%\left(\frac{\norm{\theta_0-\theta^*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\theta_0-\theta^*})}{t+1} \right)\,.
\end{align}

@@

@@

Let $P$ in \Cref{assmp:lsa}-~\ref{dist} be AS. Define $\gamma_t\eqdef U^{-1}\theta_t,\,,\gamma^*\eqdef U^{-1}\theta^*$, and suppose if $P_U$ is positive definite, then

@@

\subsection{Proof Ingredients}
\paragraph{Error Recursion and Contraction Maps:}

@@

Notice that \eqref{eq:spectralrand} places our definitions of $\rhod{P}$ and $\rhos{P}$ in \Cref{def:contract}. Here the subscripts $s$ and $d$ stand for \emph{stochastic} and \emph{deterministic}, since they are associated with the spectral radius of stochastic and deterministic matrices respectively.\\
\paragraph{Positive Definiteness:} By expanding \eqref{eq:spectralrand} we get
\begin{align*}
\EE{\norm{I-\alpha A_t}^2}=\sup_{x:\norm{x}=1} 1 -\alpha x^* (A_P+A_P^*)x +\alpha^2 \EE{A_t^* A_t},
\end{align*}
and thanks to \Cref{lm:pd} we know that when $A_P$ is positive definite it is possible to choose an $\alpha$ such that $\EE{\norm{I-\alpha A_t}}<1$.
\paragraph{Similarity Transformation:} When $A_{P^{AS}}$ is known only to be AS, it is possible that there exists no choice of $\alpha>0$ for which $\rhos{P^{AS}}>0$ and $\rhod{P^{AS}}>0$ (see \Cref{ex:matrices}). However, it is potentially possible to re-write the recursion in a different basis via a similarity transformation $\gamma_t=U^{-1 }\theta_t$, and $z_t\eqdef U^{-1} e_t$
\begin{align}\label{eq:errecsimtran}
z_t&=\underbrace{(I-\alpha J_t)}_{\text{Transformed-Matrix}} z_{t-1}+\underbrace{\alpha U^{-1}(N_t -(M_t)\ts)}_{\text{Transformed-Noise}},
where $J_t\sim P_U$ and thanks to \Cref{lm:simtran} we know that $P_U$ is PD and from \Cref{lm:pd} we know that there exists $\alpha$ such that $\EE{\norm{I-\alpha J_t}}<1$. We note here that a similarity transformation does not result in \eqref{eq:errrecsimtran} being a diagonal system. However, \Cref{lm:simtran} side steps non-diagonizablility by taking refuge in Jordon decomposition theorem in linear algebra (see Proof of \Cref{lm:simtran} in the Appendix).
\subsection{Result Discussion}
We first discuss implications of \Cref{thm:pdrate}.

### 2017-05-06 related.tex
initial version

\subsection{Reinforcement Learning}
\subsection{GTD}
\todoch{Use LiHong Li's GTD results, choice of the basis $U$ etc}

>>>>

\subsection{Reinforcement Learning}

\paragraph{RL Setting}
We are presented with data in the form of an $i.i.d$ sequence $(\phi_t,\phi'_t,r_t)\in \R^n\times\R^n\times \R$. Let $\gamma \in (0,1)$ be a fixed constant, we define $\Delta_t\eqdef \phi_t\phi_t^\top-\gamma \phi_t\phi_t^{'\top}$, $C_t=\phi_t\phi_t^\top$ and $b_t=\phi_r r_t$ the various temporal difference learning algorithms can be written as in \Cref{table:td}. In what follows, $\rho_t>0$ is an \emph{importance} sampling factor whose aim is to correct for mismatch in the (behavior) distribution with which the data was collected and the (target) distribution with respect to which one wants to learn. A factor $\rho_t=1,\,\forall t\geq 0$ will mean that no correction is required\footnote{This is known as the \emph{on-policy} case where the behavior is identical to the target. The general setting where $\rho_t>0$ is known as \emph{off-policy}.}.
\begin{comment}
\begin{table}

\begin{tabular}{|c|c|c|c|c|}\hline
Quantity& TD(0) & GTD & GTD2 & TDC \\ \hline
$b_t$& $\phi_t r_t$ & $\begin{bmatrix}\phi_t r_t\\ \mathbf{0}_{n\times 1} \end{bmatrix}$ &$\begin{bmatrix}\phi_t r_t\\ \mathbf{0}_{n\times 1} \end{bmatrix}$ & $\begin{bmatrix}\phi_t r_t\\ \phi_t r_t \end{bmatrix}$
\end{tabular}
\end{table}
\end{comment}
\begin{table}
\resizebox{\columnwidth}{!}{
\begin{tabular}{|c|c|c|c|c|}\hline
Algorithm& Update &Stepsize &Rate \\ \hline
TD(0)& $\begin{aligned}\theta_t&=\theta_{t-1}+\alpha_t (b_t -\Delta_t\theta_{t-1})\\ \thh_t&=\frac{1}{t+1}\sum_{s=0}^t \theta_t\end{aligned}$ &$\begin{aligned}\alpha_t&=O(\frac1{t})^\delta,\\ &\delta\in(1/2,1)\end{aligned}$ &$\EE{\norm{\eh_t}}\leq O(\frac{1}{\sqrt{t}})$ for $\delta\ra 1$ \\ \hline
GTD& $\begin{aligned} y_{t}&=y_{t-1}+\beta_t(\rho_t b_t -\rho_t \Delta_t \theta_{t-1}- Q_t y_{t-1})\\ \theta_t &=\theta_{t-1}+\alpha_t (\rho_t A^\top_t y_{t-1})\end{aligned}$ & $\begin{aligned}&\beta_t=\eta \alpha_t,\sum_{t\geq 0}\alpha_t=\infty, \\ &\sum_{t\geq 0}\alpha^2_t<\infty \end{aligned}$ & $e_t\ra 0$ w.p.1 \\\hline
TDC& $\begin{aligned} y_{t}&=y_{t-1}+\beta_t(\rho_t b_t -\rho_t \Delta_t \theta_{t-1}-C_t y_{t-1})\\ \theta_t &=\theta_{t-1}+\alpha_t (\rho_t b_t-\rho_t\Delta_t\theta_{t-1}+ (\Delta^\top_t-C_t)y_{t-1})\end{aligned}$ & $\begin{aligned}&\beta_t,\alpha_t,\frac{\alpha_t}{\beta_t} \ra 0,~\text{as}~t\ra \infty,\\& \sum_{t\geq 0}\alpha_t=\infty, \sum_{t\geq 0}\beta_t=\infty, \\ &\sum_{t\geq 0}\alpha^2_t<\infty, \sum_{t\geq 0}\beta^2_t<\infty \end{aligned}$ & $e_t\ra 0$ w.p.1 \\\hline
GTD& $\begin{aligned} y_{t}&=\Pi_Y \left(y_{t-1}+\alpha_t(\rho_t b_t -\rho_t \Delta_t \theta_{t-1}-Q_t y_{t-1})\right)\\ \theta_t &=\Pi_{\Theta}\left(\theta_{t-1}+\alpha_t (\rho_t A_t^\top y_{t-1})\right)\\ \thh_t&=\frac{\sum_{s=0}^t \alpha_s\theta_s}{\sum_{s=0}^t \alpha_s}\end{aligned}$ & $\alpha_t =O(\frac{1}{\sqrt{t}})$ & $\norm{A_P\eh_t-b}^2 \leq O(\frac{1}{\sqrt{t}})$ w.h.p\\\hline

\end{tabular}
}
\end{table}

1) No bias Variance Split in Korda
2) Projection, and incorrectness in GTDMP
3) TDC only SA result
3) Recent results by LiHongLi (need to say what they say clearly)

\todoch{Use LiHong Li's GTD results, choice of the basis $U$ etc}
